Final project

### 2021-12-17 final_project.Rmd
finished code

---
title: "draft_yma"
author: "Anna Ma"
date: "12/12/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(leaps)
library(readr)
```

## Data exploration

 **converted variables to per person and docs, beds, etc.**

@@

### Descriptive statistics of all varaibles
```{r statistic summary}

@@

Q: 1. do we need to group them by state or county? group by gounty gives 5000+ rows though...
  2. do we need box plot still? 
  3. again, transfer some variables to "per pop" / "per 1000 pop" ?

### Descriptive analysis

@@

boxplot(cdi$hsgrad, main = "hsgrad")
boxplot(cdi$bagrad, main = "bagrad")
boxplot(cdi$poverty, main = "poverty")
boxplot(cdi$unemp,main = "unemp")
boxplot(cdi$pcincome, main = "pcincome")
boxplot(cdi$totalinc, main = "totalinc")

### Correlation

#### Pairwise relationship

@@

library(ggcorrplot)
library(ggstatsplot)
ggstatsplot::ggcorrmat(
  data = cdi_descriptive,
  type = "parametric", # parametric for Pearson, nonparametric for Spearman's correlation
  colors = c("darkred", "white", "steelblue") # change default colors
)
```

* Here is the correlation with pairs function. More specific than the heat map above. We can observe the correlation between all the terms here
```{r}
pairs(crm_1000 ~.,data=cdi_descriptive, panel = panel.smooth, upper.panel = NULL, main = "Scatterplot Matrix")
```

#### Marginal distribution ?

```{r}
library(ggplot2)
library(ggExtra)

@@

### Distribution of outcome

@@

@@

## Model!

**Q: do we need more descriptive analysis, visualization? rate of crime for each state?**

#### Full model predictors

@@

### New tibble with transformed crm_1000 -> crm_1000_sqr
```{r}
cdi_trans = cdi_model %>% mutate(crm_1000_sqr = crm_1000^0.5) %>% select(-c(crm_1000))
```

@@

cdi_trans %>% 
  filter(crm_1000_sqr > upper + 1.5*IQR,
         crm_1000_sqr > lower - 1.5*IQR) %>% 
  dplyr::select(crm_1000_sqr) %>% 
  mutate(cty = c("Kings", "St._Loui")) %>% 
  knitr::kable(digits = 2)

# cdi %>% filter(pop == 2300664 | pop == 396685) %>% select(cty,pop)
```

@@

lm(formula = crm_1000_sqr ~ pop + pop18 + pop65 + hsgrad + bagrad + 
    poverty + unemp + pcincome + region + pdocs_1000 + pbeds_1000 + 
    density_pop, data = cdi_trans)

@@

step-wise?

@@

## Criterion Based: Adjr2 and Cp

@@

### Criterion

#### Cp

```{r}
olsrr::ols_mallows_cp(fit_back,trans_fit)
olsrr::ols_mallows_cp(fit_forward,trans_fit)
olsrr::ols_mallows_cp(fit_both,trans_fit)
olsrr::ols_mallows_cp(fit_adjr,trans_fit)
olsrr::ols_mallows_cp(fit_Cp,trans_fit)
```

#### Adj R2

```{r}
summary(fit_back) # 0.5469
summary(fit_forward) #0.5457
summary(fit_both) #0.5469
summary(fit_adjr) #0.5474
summary(fit_Cp) #0.5469
```

```{r}
library(flexmix)
BIC(fit_back) #1453.527
BIC(fit_forward) #1469.908
BIC(fit_both) #1453.527
BIC(fit_adjr) # 1458.104
BIC(fit_Cp) #1453.527
```

```{r}
back_summ =summary(fit_back) #1.331
mean(back_summ$residuals^2)

forward_summ =summary(fit_forward) #1.325
mean(forward_summ$residuals^2)

both_summ =summary(fit_both) #1.331
mean(both_summ$residuals^2)

adjr_summ =summary(fit_adjr) #1.326
mean(adjr_summ$residuals^2)

Cp_summ =summary(fit_Cp) #1.331
mean(Cp_summ$residuals^2)
```

@@

delete vif larger than 10. Non of them are greater than 10. So all of the models are good, don't have to worry about collinearity.

```{r}
par(mfrow=c(2,2))
plot(fit_back)
```

@@

### Back

#### Old Method
```{r}
# residuals vs leverage plot
plot(fit_back, which = 4)
#remove outliers
cdi_trans_without = cdi_trans[-c(1, 6, 53),]
fit_without = lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + 
    pcincome + region + pbeds_1000 + density_pop, data = cdi_trans_without)
summary(fit_without)
```

```{r}
par(mfrow=c(2,2))
plot(fit_without)
```

### Threshold: $D_{i}>\frac{4}{n} = \frac{4}{440} = 0.009$
```{r}
cook = olsrr::ols_plot_cooksd_bar(fit_back, print_plot = TRUE)
normal = cook$data$color == "normal"
outlier = cook$data$color == "outlier"
```

```{r}
outlier_df = cdi[outlier,]
```

This gives us 21 outliers. We should consider if we want to take them all out

### Build model without outlier

Here's what happens when we take them all out.

```{r}
cdi_trans_without = cdi_trans[normal,]
```

@@

```{r}
fit_without = lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + pcincome + region + pbeds_1000 + density_pop, data = cdi_trans_without)

summary(fit_without)

# Residual performance
par(mfrow=c(2,2))
olsrr::ols_plot_resid_fit(fit_without)
olsrr::ols_plot_resid_qq(fit_without)
MASS::boxcox(fit_without)

#Cp -88.45??
olsrr::ols_mallows_cp(fit_without,trans_fit)
#BIC: 1274.534
BIC(fit_without) 
#MSE
without_summ =summary(fit_without) #1.017
mean(without_summ$residuals^2)
#VIF 
performance::check_collinearity(fit_without)
```

@@

```{r}
fit_without2 = lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + pcincome + region+ pbeds_1000 + density_pop, data = cdi_trans_without)

summary(fit_without2)

# Residual performance
par(mfrow=c(2,2))
olsrr::ols_plot_resid_fit(fit_without2)
olsrr::ols_plot_resid_qq(fit_without2)
MASS::boxcox(fit_without2)

#Cp -87.5??
olsrr::ols_mallows_cp(fit_without2,trans_fit)
#BIC: 1279.154
BIC(fit_without2)
#MSE
without2_summ =summary(fit_without2) #1.013
mean(without2_summ$residuals^2)
#VIF 
performance::check_collinearity(fit_without2) 
```

@@

## Compute RMSE and adjusted R^2 by cross-validation

@@

cv_df = 
  cv_df %>%
  mutate(
    fit_back = map(train, ~lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + pcincome + region + pbeds_1000 + density_pop, data = .x)),
    trans_fit = map(train, ~lm(crm_1000_sqr ~ pop + pop18 + pop65+ hsgrad + bagrad + poverty + unemp + pcincome + region+ +pdocs_1000+pbeds_1000 + density_pop, data = .x)),
    fit_adjr = map(train, ~lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + pcincome + region+ pbeds_1000 + density_pop, data = .x)),
    fit_forward = map(train, ~lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + pcincome + region+ pbeds_1000 + density_pop, data = .x))) %>% 
  mutate(
    rmse_model2 = map2_dbl(fit_back, test, ~rmse(model = .x, data = .y)),
    rmse_model3 = map2_dbl(trans_fit, test, ~rmse(model = .x, data = .y)),
    rmse_model4 = map2_dbl(fit_adjr, test, ~rmse(model = .x, data = .y)),
    rmse_model5 = map2_dbl(fit_forward, test, ~rmse(model = .x, data = .y))) %>%
  mutate(
    mspe_model2 = map2_dbl(fit_back, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2)),
    mspe_model3 = map2_dbl(trans_fit, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2)),
    mspe_model4 = map2_dbl(fit_adjr, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2)),
    mspe_model5 = map2_dbl(fit_forward, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2))
  ) %>% 
  mutate(
    res_model2 = map(fit_back, broom::glance %>% as.data.frame),
    res_model3 = map(trans_fit, broom::glance %>% as.data.frame),
    res_model4 = map(fit_adjr, broom::glance %>% as.data.frame),
    res_model5 = map(fit_forward, broom::glance %>% as.data.frame)) %>%
  unnest(res_model2, res_model3, res_model4, res_model5) %>%
  dplyr::select(rmse_model2,rmse_model3,rmse_model4,rmse_model5,
                mspe_model2,mspe_model3,mspe_model4,mspe_model5,
                value.adj.r.squared,value.adj.r.squared1,value.adj.r.squared2,value.adj.r.squared3) %>%
  rename(adjR_model2 = value.adj.r.squared,
         adjR_model3 = value.adj.r.squared1,
         adjR_model4 = value.adj.r.squared2,
         adjR_model5 = value.adj.r.squared3)

@@

## Plot the violin plot
```{r warning=FALSE}
unnest_cd_df = cv_df %>%
  pivot_longer(rmse_model2:adjR_model5, 
               names_pattern = "(.*)(......)$", 
               names_to = c("limit", "model")) %>% 
  mutate(limit=ifelse(limit=="", "value", limit)) %>%
  pivot_wider(id_cols = model, 
              names_from = limit, 
              values_from = value, 
              names_repair = "check_unique") %>% 
  unnest(c(rmse_, adjR_)) %>%
  rename(rmse = rmse_,
         adjR = adjR_)
unnest_cd_df %>% ggplot(aes(x = model, y = rmse)) + geom_violin() 
unnest_cd_df %>% ggplot(aes(x = model, y = adjR)) + geom_violin() 
```

## backward by hand

```{r}
# No docs
step1 = update(trans_fit, . ~ . -pdocs_1000)
summary(step1)

# No pop65
step2 = update(step1, . ~ . -pop65)
summary(step2)

# No unemp
step3 = update(step2, . ~ . -unemp)
summary(step3)

step4 = update(step3, . ~ . -hsgrad)
summary(step4)

mult.fit.final = lm(crm_1000^0.5 ~ pop + pop18 + bagrad + poverty + 
    pcincome + region + pbeds_1000 + density_pop, data = cdi_model)
summary(mult.fit.final)
```

>>>>

---
title: "Final project"
author: "Anna Ma"
date: "12/17/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(leaps)
library(readr)
library(flexmix)
```

# Data exploration

@@

## Descriptive statistics of all varaibles
```{r statistic summary}

@@

## Descriptive analysis

@@

boxplot(cdi$hsgrad, main = "hsgrad")
boxplot(cdi$bagrad, main = "bagrad")
boxplot(cdi$poverty, main = "poverty")
boxplot(cdi$unemp,main = "unemp")
boxplot(cdi$pcincome, main = "pcincome")
boxplot(cdi$totalinc, main = "totalinc")
```

## Correlation

### Pairwise relationship

@@

library(ggcorrplot)

ggcorrplot(cor(cdi_descriptive), type = "upper",
   lab = TRUE)
```

Scatter plot

```{r}
pairs(crm_1000 ~.,data=cdi_descriptive, panel = panel.smooth, upper.panel = NULL, main = "Scatterplot Matrix")
```

## Marginal distribution

```{r}
library(ggplot2)
library(ggExtra)
```

@@

## Distribution of outcome

@@

## Counties with unusual rates

@@

# Model

## Full model predictors

@@

New tibble with transformed Y:
```{r}
cdi_trans = cdi_model %>% mutate(crm_1000_sqr = crm_1000^0.5) %>% select(-c(crm_1000))
```

@@

cdi_trans %>% 
  filter(crm_1000_sqr > upper + 1.5*IQR,
         crm_1000_sqr > lower - 1.5*IQR) %>% 
  dplyr::select(crm_1000_sqr) %>% 
  mutate(cty = c("Kings", "St._Loui")) %>% 
  knitr::kable(digits = 2)
```

## Model Selection

### Step-wise approach

@@

model: crm_1000_sqr ~ pop + pop18 + pop65 + hsgrad + bagrad + 
    poverty + unemp + pcincome + region + pdocs_1000 + pbeds_1000 + 
    density_pop, data = cdi_trans)

@@

step-wise

@@

### Criterion Based: Adjr2 and Cp

@@

@@

Delete predictors when vif larger than 10. Non of them are greater than 10. So all of the models are good, don't have to worry about collinearity.

@@

```{r}
olsrr::ols_plot_cooksd_bar(fit_back, print_plot = TRUE)
#remove observations whose cook's D > 0.5
cdi_trans_without = cdi_trans[-c(1, 6),]
```

## Build model without outlier

@@

```{r}
fit_back_without = lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + pcincome + region + pbeds_1000 + density_pop, data = cdi_trans_without)

summary(fit_back_without)

# Residual performance
par(mfrow=c(2,2))
olsrr::ols_plot_resid_fit(fit_back_without)
olsrr::ols_plot_resid_qq(fit_back_without)
```

Build the without model with **trans_fit** (full model)

```{r}
trans_fit_without = lm(crm_1000_sqr ~., data = cdi_trans_without)

summary(trans_fit_without)

# Residual performance
par(mfrow=c(2,2))
olsrr::ols_plot_resid_fit(trans_fit_without)
olsrr::ols_plot_resid_qq(trans_fit_without)
```

@@

```{r}
fit_adjr_without = lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + pcincome + region+ pbeds_1000 + density_pop, data = cdi_trans_without)

summary(fit_adjr_without)

# Residual performance
par(mfrow=c(2,2))
olsrr::ols_plot_resid_fit(fit_adjr_without)
olsrr::ols_plot_resid_qq(fit_adjr_without)
```

## Criteria

#### Cp for models fit without outlier

```{r}
olsrr::ols_mallows_cp(fit_back_without,trans_fit) #-14.65
olsrr::ols_mallows_cp(trans_fit_without,trans_fit)# -10.59
olsrr::ols_mallows_cp(fit_adjr_without,trans_fit) #-14.49
```

#### AIC

```{r}
AIC(fit_back_without) # 1371.373
AIC(trans_fit_without) #1375.258
AIC(fit_adjr_without) # 1371.363
```

```{r}
BIC(fit_back_without) #1424.441
BIC(trans_fit_without) #1440,573
BIC(fit_adjr_without) # 1428.514
```

```{r}
back_without_summ =summary(fit_back_without) #1.2633
mean(back_without_summ$residuals^2)

trans_without_summ =summary(trans_fit_without) #1.2572
mean(trans_without_summ$residuals^2)

adjr_without_summ =summary(fit_adjr_without) #1.2575
mean(adjr_without_summ$residuals^2)
```

@@

## Compute RMSE, MSPE, adjusted R^2 by cross-validation

@@

cv_df = 
  cv_df %>%
  mutate(
    fit_back = map(train, ~lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + pcincome + region + pbeds_1000 + density_pop, data = .x)),
    trans_fit = map(train, ~lm(crm_1000_sqr ~ pop + pop18 + pop65+ hsgrad + bagrad + poverty + unemp + pcincome + region+ +pdocs_1000+pbeds_1000 + density_pop, data = .x)),
    fit_adjr = map(train, ~lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + pcincome + region+ pbeds_1000 + density_pop, data = .x))) %>% 
  mutate(
    rmse_back = map2_dbl(fit_back, test, ~rmse(model = .x, data = .y)),
    rmse_full = map2_dbl(trans_fit, test, ~rmse(model = .x, data = .y)),
    rmse_adjr = map2_dbl(fit_adjr, test, ~rmse(model = .x, data = .y))) %>%
  mutate(
    mspe_back = map2_dbl(fit_back, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2)),
    mspe_full = map2_dbl(trans_fit, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2)),
    mspe_adjr = map2_dbl(fit_adjr, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2))) %>% 
  mutate(
    res_back = map(fit_back, broom::glance %>% as.data.frame),
    res_full = map(trans_fit, broom::glance %>% as.data.frame),
    res_adjr = map(fit_adjr, broom::glance %>% as.data.frame)) %>%
  unnest(res_back, res_full, res_adjr) %>%
  dplyr::select(rmse_back,rmse_full,rmse_adjr,
                mspe_back,mspe_full,mspe_adjr,
                value.adj.r.squared,value.adj.r.squared1,value.adj.r.squared2) %>%
  rename(adjR_back = value.adj.r.squared,
         adjR_full = value.adj.r.squared1,
         adjR_adjr = value.adj.r.squared2)

@@

## Plot the violin plot
```{r warning=FALSE}
unnest_cd_df = cv_df %>%
  pivot_longer(rmse_back:adjR_adjr, 
               names_pattern = "(.*)(....)$", 
               names_to = c("limit", "model")) %>% 
  mutate(limit=ifelse(limit=="", "value", limit)) %>%
  pivot_wider(id_cols = model, 
              names_from = limit, 
              values_from = value, 
              names_repair = "check_unique") %>% 
  unnest(c(rmse_, adjR_)) %>%
  rename(rmse = rmse_,
         adjR = adjR_)
unnest_cd_df %>% ggplot(aes(x = model, y = rmse)) + geom_violin() 
unnest_cd_df %>% ggplot(aes(x = model, y = adjR)) + geom_violin() 
```
